### 2021-10-08 Output/Figures/E4/coverage_posteriormean.tex
E6 on dcc

% latex table generated in R 4.0.2 by xtable 1.8-4 package
% Wed Oct 06 16:48:58 2021
\begin{table}[ht]
\centering
\begin{tabular}{ccccc}
  \toprule
 & var & cover & mean\_overall & mean\_sd \\ 
  \midrule
1 & beta & 0.895 & 0.0025 & 0.000255 \\ 
  2 & gamma & 0.902 &    1 & 0.142 \\ 
  3 & R0 & 0.910 & 2.54 & 0.194 \\ 
   \bottomrule
\end{tabular}
\caption{Nomial coverage rate and distribution of posterior means among 2000 simulations. The true value of the parameters are (beta, gamma, R_0) = (0.0025, 1, 2.5)} 
\label{tab:cov}
\end{table}